---
title: "State Regression + Polls"
author: "Emma Landry"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(epidemia)
library(dplyr)
library(here)
library(readr)
library(readxl)
library(rstan)
library(ggplot2)
library(GGally)
library(reshape2) #for melt
library(urbnmapr)
library(tidyverse)
library(rstanarm)
```

```{r}
short_state <- c("AL","AZ", "AR", "CA","CO","CT","DE","FL","GA","ID","IL","IN","IA","KS","KY",
                 "LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND",
                 "OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY") 
```

Columns Needed:
-Over 18 total pop
-Over 18 population hesitating
-don't like vaccines
-dont trust covid 19 vaccines
-don't trust the government
-other people need it more right now (could be negatively correlated with trump voting)

```{r}
zeros <- rep(0, 48)
vax <- data.frame("STATE" = short_state,
                  "TOT_POP" = zeros, 
                  "TOT_HESIT" = zeros,
                  "NO_LIKE" = zeros,
                  "NO_VAX_TRUST"= zeros,
                  "NO_GOV_TRUST"= zeros,
                  "OTHERS_NEED"= zeros)
```


```{r}
for (i in seq(1:48)){
  df1 <- read_excel(here::here("data", "vaccination_hesitancy", "health5_week22.xlsx"), sheet = short_state[i])
  df2 <-read_excel(here::here("data", "vaccination_hesitancy", "health6_week22.xlsx"), sheet = short_state[i])
  vax$TOT_POP[i] <- df1[8,2]
  vax$TOT_HESIT[i]<- df2[7,2]
  vax$NO_LIKE[i]<- df2[7,6]
  vax$NO_VAX_TRUST[i] <- df2[7,11]
  vax$NO_GOV_TRUST[i] <- df2[7,12]
  vax$OTHERS_NEED[i] <- df2[7,9]
}
```

```{r}
vax$TOT_POP <- as.numeric(vax$TOT_POP)
vax$TOT_HESIT <- as.numeric(vax$TOT_HESIT)
vax$NO_LIKE <- as.numeric(vax$NO_LIKE)
vax$NO_VAX_TRUST<- as.numeric(vax$NO_VAX_TRUST)
vax$NO_GOV_TRUST <- as.numeric(vax$NO_GOV_TRUST)
vax$OTHERS_NEED <- as.numeric(vax$OTHERS_NEED)
```

```{r}
vax_prop <- data.frame("STATE"= vax$STATE,
                       "TOT_POP"= vax$TOT_POP,
                       "HESIT" = vax$TOT_HESIT/ vax$TOT_POP,
                       "NO_LIKE" = vax$NO_LIKE/vax$TOT_HESIT,
                       "NO_VAX_TRUST"= vax$NO_VAX_TRUST/vax$TOT_HESIT,
                       "NO_GOV_TRUST"= vax$NO_GOV_TRUST/vax$TOT_HESIT,
                       "OTHERS_NEED"= vax$OTHERS_NEED/vax$TOT_HESIT
                       )
```

```{r}
elec <- read.csv(here::here("data", "election", "1976-2020-president.csv"))
elec <- subset(elec, ((elec$year==2020 | elec$year==2016)& elec$party_simplified=="REPUBLICAN"))
elec <- subset(elec, (elec$state!= "DISTRICT OF COLUMBIA" & elec$state != "ALASKA" & elec$state!= "HAWAII" & elec$writein == FALSE) )
```

```{r}
prop_2020 <- subset(elec, elec$year==2020)$candidatevotes /subset(elec, elec$year==2020)$totalvotes
prop_2016 <- subset(elec, elec$year==2016)$candidatevotes /subset(elec, elec$year==2016)$totalvotes
change <- (prop_2020-prop_2016)
```


```{r}
trump_data <- data.frame("STATE"= subset(elec, elec$year==2020)$state_po,
                         "VOTES_2020" = prop_2020,
                         "CHANGE_2020" = change)
```

```{r}
path <- "/Volumes/evl17/home/Msc/Outputs/rt_medians/state1/"
path <- paste0(path, "alabama_3395374[1].pbs_medians.rds")
```

```{r}
long_state <- c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut",
                "delaware", "florida", "georgia", "idaho", "illinois","indiana","iowa", "kansas",
                "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan",
                "minnesota", "mississippi", "missouri","montana", "nebraska", "nevada","new_hampshire",
                "new_jersey", "new_mexico", "new_york", "north_carolina", "north_dakota", "ohio",
                "oklahoma", "oregon", "pennsylvania", "rhode_island", "south_carolina",
                "south_dakota", "tennessee", "texas", "utah", "vermont", "virginia", "washington",
                "west_virginia","wisconsin", "wyoming")
```

```{r}
file_list <- list.files(path= "/Volumes/evl17/home/Msc/Outputs/rt_medians/state1")
file_list
```

```{r}
epifit_sc <- readRDS("/Volumes/evl17/home/Msc/Outputs/epidemia_fits/state1/rhode_island_3395374[38].pbs.rds")
```

```{r}
rt <- posterior_rt(epifit_sc$fit)
times<- rt$time
```

```{r}
med_df <- data.frame(matrix(ncol = 379, nrow = 49))
colnames(med_df)<- times
med_df$state <- NA
med_df <- med_df %>%
  select(state, everything())
```


```{r}
count <- 1
  
for (i in seq(1:49)){
  this.long <- long_state[i]
  
  path <- paste0("/Volumes/evl17/home/Msc/Outputs/rt_medians/state1/", file_list[i])
  medians <- read.csv(path)
  medians_end <- tail(medians, 379)
  vec <- c(this.long, as.numeric(medians_end$x))

  med_df[count,] <- vec
  count <- count +1
}
```


```{r}
topline_elec <- read.csv(here::here("data", "election", "presidential_state_toplines_2020.csv"))
```

```{r}
topline_elec_cut <- topline_elec[,-c(1,2,3,7,9,10,11,12,13)]
topline_elec_cut <- topline_elec_cut[,-(9:28)]
topline_elec_cut <- topline_elec_cut[,-7]
topline_elec_cut <- topline_elec_cut %>% arrange(-row_number())
```


```{r}
topline_elec_cut <- subset(topline_elec_cut, (state!= "ME-2" & state!= "ME-1" & state != "Hawaii"&state!= "NE-1" &state!= "NE-2" &state!= "NE-3" & state!= "District of Columbia" ))
```

```{r}
nrow(subset(topline_elec_cut, modeldate == "11/3/2020"))
```

```{r}
dates_elec <-unique(topline_elec_cut$modeldate)
```

```{r}
poll_forecasts_trump <- data.frame(matrix(ncol = length(dates_elec), nrow = 49))
colnames(poll_forecasts_trump)<- dates_elec
poll_forecasts_trump$state <- long_state
poll_forecasts_trump <- poll_forecasts_trump%>%
  select(state, everything())
```


```{r}
for (i in (1:length(dates_elec))){
  time <- dates_elec[i]
  mini_df <- subset(topline_elec_cut, modeldate == time)
  poll_forecasts_trump[,i+1]<- mini_df$voteshare_inc
}
```

```{r}
poll_forecasts_trump <- subset(poll_forecasts_trump, state != "alaska")
poll_forecasts_trump$`11/4/2020`<- trump_data$VOTES_2020*100
```

```{r}
med_df_cut <- med_df[,c(1,77:380)]
med_df_cut<- subset(med_df_cut, state!= "alaska")
med_df_cut[,2:305] <- lapply(med_df_cut[,2:305],as.numeric)
```

```{r}
ncol(med_df[,2:15])
```
```{r}
trump_coeffs <- c()
hes_coeffs <- c()
```


```{r}
ind1 <- 2
ind2 <- ind1+13
for (i in seq(1:16)){
  if (i<=10){ 
   means_rt <-  rowMeans(med_df_cut[,ind1:ind2])
   means_forecast <- rowMeans(poll_forecasts_trump[,ind1:ind2])
   matrix <- data.frame("Medians"= means_rt,
                    "Trump"= means_forecast,
                    "Hes" = vax_prop$HESIT)
  }
  else{ #We no longer use poll data just results
    means_rt <-  rowMeans(med_df_cut[,ind1:ind2])
    forecast <- poll_forecasts_trump$`11/4/2020`
    matrix <- data.frame("Medians"= means_rt,
                    "Trump"= forecast,
                    "Hes" = vax_prop$HESIT)
  }
  
  model <- lm(Medians ~ Trump + Hes, data= matrix) 
  trump_coeffs <- c(trump_coeffs, model$coefficients[2])
  hes_coeffs <- c(hes_coeffs, model$coefficients[3])
  
  ind1<- ind2+1
  ind2 <- ind1+13

}

```
```{r}
plot(hes_coeffs)
```
Until index 10- general trend is negative correlation between vaccination hesitancy and Rt. Higher vaccination hesitancy, lower Rt- not very logical. Note: using vaccination hesitancy from mid-January. This means in the months leading up to election, higher vaccination hesitancy in January is actually counter informative, i.e. the states that will be most hesitant in January, actually had lower Rt in the fall.

```{r}
plot(trump_coeffs)
```
Until index 10 (election day), trump voting is positively correlated with higher Rt. After that date, where we only use the results from November 4th.


We can observe a random date in fall, and one in december

```{r}
plot_df_sept28 <- data.frame("TrumpSupport"= poll_forecasts_trump[,121],
                       "RtMedian" = med_df_cut[,121],
                       "Hesitancy" = vax_prop$HESIT,
                       "State"= str_to_title(gsub("_", " ", med_df_cut$state)))

plot_df_dec16 <- data.frame("TrumpSupport"= poll_forecasts_trump[,158],
                       "RtMedian" = med_df_cut[,200],
                       "Hesitancy" = vax_prop$HESIT,
                       "State"= str_to_title(gsub("_", " ", med_df_cut$state)))

```

```{r}
sp1 <- ggplot(data= plot_df_sept28, aes(x=TrumpSupport,y=RtMedian,label = State))+
  geom_point()+
  geom_text()
sp1

sp2 <- ggplot(data= plot_df_dec16, aes(x=TrumpSupport,y=RtMedian,label = State))+
  geom_point()+
  geom_text()
sp2
```



```{r}
sp3 <- ggplot(data= plot_df_sept28, aes(x=Hesitancy,y=RtMedian,label = State))+
  geom_point()+
  geom_text()
sp3

sp4 <- ggplot(data= plot_df_dec16, aes(x=Hesitancy,y=RtMedian,label = State))+
  geom_point()+
  geom_text()
sp4
```




