---
title: "State Regression + Polls"
author: "Emma Landry"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(epidemia)
library(dplyr)
library(here)
library(readr)
library(readxl)
library(rstan)
library(ggplot2)
library(GGally)
library(reshape2) #for melt
library(urbnmapr)
library(tidyverse)
library(rstanarm)
```

```{r}
short_state <- c("AL","AZ", "AR", "CA","CO","CT","DE","FL","GA","ID","IL","IN","IA","KS","KY",
                 "LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND",
                 "OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY") 
```

Columns Needed:
-Over 18 total pop
-Over 18 population hesitating
-don't like vaccines
-dont trust covid 19 vaccines
-don't trust the government
-other people need it more right now (could be negatively correlated with trump voting)

```{r}
zeros <- rep(0, 48)
vax <- data.frame("STATE" = short_state,
                  "TOT_POP" = zeros, 
                  "TOT_HESIT" = zeros,
                  "NO_LIKE" = zeros,
                  "NO_VAX_TRUST"= zeros,
                  "NO_GOV_TRUST"= zeros,
                  "OTHERS_NEED"= zeros)
```


```{r}
for (i in seq(1:48)){
  df1 <- read_excel(here::here("data", "vaccination_hesitancy", "health5_week22.xlsx"), sheet = short_state[i])
  df2 <-read_excel(here::here("data", "vaccination_hesitancy", "health6_week22.xlsx"), sheet = short_state[i])
  vax$TOT_POP[i] <- df1[8,2]
  vax$TOT_HESIT[i]<- df2[7,2]
  vax$NO_LIKE[i]<- df2[7,6]
  vax$NO_VAX_TRUST[i] <- df2[7,11]
  vax$NO_GOV_TRUST[i] <- df2[7,12]
  vax$OTHERS_NEED[i] <- df2[7,9]
}
```

```{r}
vax$TOT_POP <- as.numeric(vax$TOT_POP)
vax$TOT_HESIT <- as.numeric(vax$TOT_HESIT)
vax$NO_LIKE <- as.numeric(vax$NO_LIKE)
vax$NO_VAX_TRUST<- as.numeric(vax$NO_VAX_TRUST)
vax$NO_GOV_TRUST <- as.numeric(vax$NO_GOV_TRUST)
vax$OTHERS_NEED <- as.numeric(vax$OTHERS_NEED)
```

```{r}
vax_prop <- data.frame("STATE"= vax$STATE,
                       "TOT_POP"= vax$TOT_POP,
                       "HESIT" = vax$TOT_HESIT/ vax$TOT_POP,
                       "NO_LIKE" = vax$NO_LIKE/vax$TOT_HESIT,
                       "NO_VAX_TRUST"= vax$NO_VAX_TRUST/vax$TOT_HESIT,
                       "NO_GOV_TRUST"= vax$NO_GOV_TRUST/vax$TOT_HESIT,
                       "OTHERS_NEED"= vax$OTHERS_NEED/vax$TOT_HESIT
                       )
```

```{r}
elec <- read.csv(here::here("data", "election", "1976-2020-president.csv"))
elec <- subset(elec, ((elec$year==2020 | elec$year==2016)& elec$party_simplified=="REPUBLICAN"))
elec <- subset(elec, (elec$state!= "DISTRICT OF COLUMBIA" & elec$state != "ALASKA" & elec$state!= "HAWAII" & elec$writein == FALSE) )
```

```{r}
prop_2020 <- subset(elec, elec$year==2020)$candidatevotes /subset(elec, elec$year==2020)$totalvotes
prop_2016 <- subset(elec, elec$year==2016)$candidatevotes /subset(elec, elec$year==2016)$totalvotes
change <- (prop_2020-prop_2016)
```


```{r}
trump_data <- data.frame("STATE"= subset(elec, elec$year==2020)$state_po,
                         "VOTES_2020" = prop_2020,
                         "CHANGE_2020" = change)
```

```{r}
path <- "/Volumes/evl17/home/Msc/Outputs/rt_medians/state1/"
path <- paste0(path, "alabama_3395374[1].pbs_medians.rds")
```

```{r}
long_state <- c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut",
                "delaware", "florida", "georgia", "idaho", "illinois","indiana","iowa", "kansas",
                "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan",
                "minnesota", "mississippi", "missouri","montana", "nebraska", "nevada","new_hampshire",
                "new_jersey", "new_mexico", "new_york", "north_carolina", "north_dakota", "ohio",
                "oklahoma", "oregon", "pennsylvania", "rhode_island", "south_carolina",
                "south_dakota", "tennessee", "texas", "utah", "vermont", "virginia", "washington",
                "west_virginia","wisconsin", "wyoming")
```

```{r}
file_list <- list.files(path= "/Volumes/evl17/home/Msc/Outputs/rt_medians/state1")
file_list
```

```{r}
epifit_sc <- readRDS("/Volumes/evl17/home/Msc/Outputs/epidemia_fits/state1/rhode_island_3395374[38].pbs.rds")
```

```{r}
rt <- posterior_rt(epifit_sc$fit)
times<- rt$time
```

```{r}
med_df <- data.frame(matrix(ncol = 379, nrow = 49))
colnames(med_df)<- times
med_df$state <- NA
med_df <- med_df %>%
  select(state, everything())
```


```{r}
count <- 1
  
for (i in seq(1:49)){
  this.long <- long_state[i]
  
  path <- paste0("/Volumes/evl17/home/Msc/Outputs/rt_medians/state1/", file_list[i])
  medians <- read.csv(path)
  medians_end <- tail(medians, 379)
  vec <- c(this.long, as.numeric(medians_end$x))

  med_df[count,] <- vec
  count <- count +1
}
```


```{r}
topline_elec <- read.csv(here::here("data", "election", "presidential_state_toplines_2020.csv"))
```

```{r}
topline_elec_cut <- topline_elec[,-c(1,2,3,7,9,10,11,12,13)]
topline_elec_cut <- topline_elec_cut[,-(9:28)]
topline_elec_cut <- topline_elec_cut[,-7]
topline_elec_cut <- topline_elec_cut %>% arrange(-row_number())
```


```{r}
topline_elec_cut <- subset(topline_elec_cut, (state!= "ME-2" & state!= "ME-1" & state != "Hawaii"&state!= "NE-1" &state!= "NE-2" &state!= "NE-3" & state!= "District of Columbia" ))
```

```{r}
nrow(subset(topline_elec_cut, modeldate == "11/3/2020"))
```

```{r}
dates_elec <-unique(topline_elec_cut$modeldate)
```

```{r}
poll_forecasts_trump <- data.frame(matrix(ncol = length(dates_elec), nrow = 49))
colnames(poll_forecasts_trump)<- dates_elec
poll_forecasts_trump$state <- long_state
poll_forecasts_trump <- poll_forecasts_trump%>%
  select(state, everything())
```


```{r}
for (i in (1:length(dates_elec))){
  time <- dates_elec[i]
  mini_df <- subset(topline_elec_cut, modeldate == time)
  poll_forecasts_trump[,i+1]<- mini_df$voteshare_inc
}
```

```{r}
poll_forecasts_trump <- subset(poll_forecasts_trump, state != "alaska")
poll_forecasts_trump$`11/4/2020`<- trump_data$VOTES_2020*100
```

```{r}
correcting_bias <- rowMeans(poll_forecasts_trump[,2:157])
correcting_bias
poll_forecasts_trump$`11/4/2020`

mean(poll_forecasts_trump$`11/4/2020`-correcting_bias)
poll_forecasts_trump$`11/3/2020`
mean(poll_forecasts_trump$`11/4/2020`- poll_forecasts_trump$`11/3/2020`)
poll_forecasts_trump$state
```


```{r}
correction <- poll_forecasts_trump$`11/4/2020`/ poll_forecasts_trump$`11/3/2020`
corrected_poll_forecast <- poll_forecasts_trump
corrected_poll_forecast[,2:157] <-  poll_forecasts_trump[,2:157]* correction
mean(poll_forecasts_trump$`11/4/2020`-rowMeans(corrected_poll_forecast[,2:157]))
mean(poll_forecasts_trump$`11/3/2020`-rowMeans(corrected_poll_forecast[,2:157]))

```


```{r}
med_df_cut <- med_df[,c(1,77:380)]
med_df_cut<- subset(med_df_cut, state!= "alaska")
med_df_cut[,2:305] <- lapply(med_df_cut[,2:305],as.numeric)
```

```{r}
ncol(med_df[,2:15])
```

```{r}
trump_coeffs <- c()
trump_10 <- c()
trump_90 <- c()

hes_coeffs <- c()
hes_10 <- c()
hes_90 <- c()

r2_means <- c()

ind1 <- 2
ind2 <- ind1+13
for (i in seq(1:16)){
  if (i<=10){ 
   means_rt <-  rowMeans(med_df_cut[,ind1:ind2])
   means_forecast <- rowMeans(corrected_poll_forecast[,ind1:ind2])
   matrix <- data.frame("Medians"= means_rt,
                    "Trump"= means_forecast,
                    "Hes" = vax_prop$HESIT)
  }
  else{ #We no longer use poll data just results
    means_rt <-  rowMeans(med_df_cut[,ind1:ind2])
    forecast <- corrected_poll_forecast$`11/4/2020`
    matrix <- data.frame("Medians"= means_rt,
                    "Trump"= forecast,
                    "Hes" = vax_prop$HESIT)
  }
  
  model <- stan_lm(Medians ~ Trump + Hes, data= matrix, prior = R2(0.5, what = "mean")) 
  model_summary<- summary(model)
  
  trump_coeffs <- c(trump_coeffs, model_summary[2,5])
  trump_10 <- c(trump_10, model_summary[2,4])
  trump_90 <- c(trump_90, model_summary[2,6])
  
  hes_coeffs <- c(hes_coeffs, model_summary[3,5])
  hes_10 <- c(hes_10, model_summary[3,4])
  hes_90 <- c(hes_90, model_summary[3,6])
  
  r2_means <- c(r2_means, model_summary[6,5])
  
  ind1<- ind2+1
  ind2 <- ind1+13

}

```
```{r}
ind1 <- 2
ind2 <- ind1+13
for (i in seq(1:16)){
  ind1<- ind2+1
  ind2 <- ind1+13

}
```



```{r}
model_summary[1:3,]
model_summary[3,]
model_summary[3,6]
hes_90
```

```{r}
model_summary[2,1]
model_summary[2,5]
model_summary[1:2,]
```

```{r}
time_varying_state <- data.frame("Week"= seq(1:16), "Hesitancy"= hes_coeffs, "Trump"= trump_coeffs,
                                 "Hesitancy10"= hes_10, "Hesitancy90"= hes_90,
                                 "Trump10"= trump_10, "Trump90"= trump_90)
```


```{r}
ggplot(data=time_varying_state, aes (x= Week))+
  geom_point(aes( y= Hesitancy, color = "Hesitancy"), size = 2)+
  geom_line(aes( y= Hesitancy), color = "#F8766D", size = 0.5)+
  geom_ribbon(aes(ymin = Hesitancy10, ymax = Hesitancy90), fill = "#F8766D", alpha = 0.2)+
  geom_point(aes(x= Week, y = Trump*150, color = "Trump Support"), size = 2)+
  geom_line(aes(x= Week, y = Trump*150), color = "#00BFC4", size = 0.5)+
  geom_ribbon(aes(ymin = Trump10*150, ymax = Trump90*150), fill = "#00BFC4", alpha = 0.2)+
  geom_hline(yintercept = 0, linetype = "dashed", color = "black")+
  scale_y_continuous(sec.axis = sec_axis( trans=~./150, name="Trump Support"))+
  scale_color_manual(values = c("Hesitancy"= "#F8766D", "Trump Support"="#00BFC4"))+
  labs(color = "",
       x = "2-Week Period")+
  theme_bw()+
  theme(
    legend.position = c(.1, .15),
    legend.background = element_rect(fill = "transparent"),
    legend.key = element_rect(fill = "transparent", colour = "transparent"))+
  scale_x_continuous(breaks=seq(1,16,by= 2),limits = c(1,16), expand = c(0, 0))
ggsave("state_time_varying_coefs.pdf", height = 5, scale =1)
```


Until index 10- general trend is negative correlation between vaccination hesitancy and Rt. Higher vaccination hesitancy, lower Rt- not very logical. Note: using vaccination hesitancy from mid-January. This means in the months leading up to election, higher vaccination hesitancy in January is actually counter informative, i.e. the states that will be most hesitant in January, actually had lower Rt in the fall.

Until index 10 (election day), trump voting is positively correlated with higher Rt. After that date, where we only use the results from November 4th.


We can observe a random date in fall, and one in december

```{r}
plot_df_sept28 <- data.frame("TrumpSupport"= corrected_poll_forecast[,121],
                       "RtMedian" = med_df_cut[,121],
                       "Hesitancy" = vax_prop$HESIT*100,
                       "State"= str_to_title(gsub("_", " ", med_df_cut$state)),
                       "ShortState"= vax$STATE)

plot_df_nov4 <- data.frame("TrumpSupport"= corrected_poll_forecast[,158],
                       "RtMedian" = med_df_cut[,158],
                       "Hesitancy" = vax_prop$HESIT*100,
                       "State"= str_to_title(gsub("_", " ", med_df_cut$state)),
                       "ShortState"= vax$STATE)


plot_df_dec16 <- data.frame("TrumpSupport"= corrected_poll_forecast[,158],
                       "RtMedian" = med_df_cut[,200],
                       "Hesitancy" = vax_prop$HESIT*100,
                       "State"= str_to_title(gsub("_", " ", med_df_cut$state)),
                       "ShortState"= vax$STATE)

```


```{r}
lm1<- stan_lm (RtMedian ~ TrumpSupport, data = plot_df_nov4, prior = R2(0.5, what = "mean"))
```

```{r}
int1<- summary(lm1)[1,5]
int10perc <- summary(lm1)[1,4]
int90perc <- summary(lm1)[1,6]

pred1 <- summary(lm1)[2,5]
pred10perc <- summary(lm1)[2,4]
pred90perc <- summary(lm1)[2,6]
```

```{r}
x<- seq(25,75, by = 0.1)
y<- int1+ x*pred1
y10perc <- int10perc+ x*pred10perc
y90perc <- int90perc+x*pred90perc
```


```{r}
trend_line_nov4<- data.frame(matrix(nrow=501))
trend_line_nov4$x <- x
trend_line_nov4$y<- y
trend_line_nov4$y10perc <- y10perc
trend_line_nov4$y90perc<- y90perc
trend_line_nov4<- trend_line_nov4[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_nov4, aes(x=TrumpSupport,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_nov4, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_nov4, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Trump Support (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(25,75), expand = c(0, 0)) +
  ggtitle("November 3rd, 2020")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("trumpreg_nov4.pdf", height = 5, scale =1)
```

```{r}
lm2<- stan_lm (RtMedian ~ Hesitancy, data = plot_df_nov4, prior = R2(0.5, what = "mean"))
```

```{r}
int2<- summary(lm2)[1,5]
int10perc2 <- summary(lm2)[1,4]
int90perc2 <- summary(lm2)[1,6]

pred2 <- summary(lm2)[2,5]
pred10perc2 <- summary(lm2)[2,4]
pred90perc2 <- summary(lm2)[2,6]
```

```{r}
x2<- seq(30,65, by = 0.1)
y2<- int2+ x2*pred2
y10perc2 <- int10perc2+ x2*pred10perc2
y90perc2 <- int90perc2+x2*pred90perc2
```


```{r}
trend_line_nov4_2<- data.frame(matrix(nrow=351))
trend_line_nov4_2$x <- x2
trend_line_nov4_2$y<- y2
trend_line_nov4_2$y10perc <- y10perc2
trend_line_nov4_2$y90perc<- y90perc2
trend_line_nov4_2<- trend_line_nov4_2[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_nov4, aes(x=Hesitancy,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_nov4_2, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_nov4_2, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Vaccine Hesitancy (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(30,65), expand = c(0, 0)) +
  ggtitle("November 3rd, 2020")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("hesitreg_nov4.pdf", height = 5, scale =1)
```
```{r}
lm1_sept28<- stan_lm (RtMedian ~ TrumpSupport, data = plot_df_sept28, prior = R2(0.5, what = "mean"))
```

```{r}
int1_sept28<- summary(lm1_sept28)[1,5]
int10perc_sept28 <- summary(lm1_sept28)[1,4]
int90perc_sept28 <- summary(lm1_sept28)[1,6]

pred1_sept28 <- summary(lm1_sept28)[2,5]
pred10perc_sept28 <- summary(lm1_sept28)[2,4]
pred90perc_sept28 <- summary(lm1_sept28)[2,6]
```

```{r}
x_sept28<- seq(30,75, by = 0.1)
y_sept28<- int1_sept28+ x_sept28*pred1_sept28
y10perc_sept28 <- int10perc_sept28+ x_sept28*pred10perc_sept28
y90perc_sept28 <- int90perc+x_sept28*pred90perc_sept28
```


```{r}
trend_line_sept28<- data.frame(matrix(nrow=451))
trend_line_sept28$x <- x_sept28
trend_line_sept28$y<- y_sept28
trend_line_sept28$y10perc <- y10perc_sept28
trend_line_sept28$y90perc<- y90perc_sept28
trend_line_sept28<- trend_line_sept28[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_sept28, aes(x=TrumpSupport,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_sept28, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_sept28, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Trump Support (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(30,75), expand = c(0, 0)) +
  ggtitle("September 28th, 2020")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("trumpreg_sept28.pdf", height = 5, scale =1)
```
```{r}
lm2_sept28<- stan_lm (RtMedian ~ Hesitancy, data = plot_df_sept28, prior = R2(0.5, what = "mean"))
```

```{r}
int2_sept28<- summary(lm2_sept28)[1,5]
int10perc2_sept28 <- summary(lm2_sept28)[1,4]
int90perc2_sept28 <- summary(lm2_sept28)[1,6]

pred2_sept28 <- summary(lm2_sept28)[2,5]
pred10perc2_sept28 <- summary(lm2_sept28)[2,4]
pred90perc2_sept28 <- summary(lm2_sept28)[2,6]
```

```{r}
x2_sept28<- seq(30,61, by = 0.1)
y2_sept28<- int2_sept28+ x2_sept28*pred2_sept28
y10perc2_sept28 <- int10perc2_sept28+ x2_sept28*pred10perc2_sept28
y90perc2_sept28 <- int90perc2_sept28+x2_sept28*pred90perc2_sept28
```


```{r}
trend_line_sept28_2<- data.frame(matrix(nrow=311))
trend_line_sept28_2$x <- x2_sept28
trend_line_sept28_2$y<- y2_sept28
trend_line_sept28_2$y10perc <- y10perc2_sept28
trend_line_sept28_2$y90perc<- y90perc2_sept28
trend_line_sept28_2<- trend_line_sept28_2[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_sept28, aes(x=Hesitancy,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_sept28_2, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_sept28_2, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Vaccine Hesitancy (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(30,61), expand = c(0, 0)) +
  ggtitle("September 28th, 2020")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("hesitreg_sept28.pdf", height = 5, scale =1)
```
```{r}
lm1_dec16<- stan_lm (RtMedian ~ TrumpSupport, data = plot_df_dec16, prior = R2(0.5, what = "mean"))
```

```{r}
int1_dec16<- summary(lm1_dec16)[1,5]
int10perc_dec16 <- summary(lm1_dec16)[1,4]
int90perc_dec16 <- summary(lm1_dec16)[1,6]

pred1_dec16 <- summary(lm1_dec16)[2,5]
pred10perc_dec16 <- summary(lm1_dec16)[2,4]
pred90perc_dec16 <- summary(lm1_dec16)[2,6]
```

```{r}
x_dec16<- seq(29,71, by = 0.1)
y_dec16<- int1_dec16+ x_dec16*pred1_dec16
y10perc_dec16 <- int10perc_dec16+ x_dec16*pred10perc_dec16
y90perc_dec16 <- int90perc_dec16+x_dec16*pred90perc_dec16
```


```{r}
trend_line_dec16<- data.frame(matrix(nrow=421))
trend_line_dec16$x <- x_dec16
trend_line_dec16$y<- y_dec16
trend_line_dec16$y10perc <- y10perc_dec16
trend_line_dec16$y90perc<- y90perc_dec16
trend_line_dec16<- trend_line_dec16[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_dec16, aes(x=TrumpSupport,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_dec16, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_dec16, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Trump Support (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(29,71), expand = c(0, 0)) +
  ggtitle("December 16th, 2020")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("trumpreg_dec16.pdf", height = 5, scale =1)
```
```{r}
lm2_dec16<- stan_lm (RtMedian ~ Hesitancy, data = plot_df_dec16, prior = R2(0.5, what = "mean"))
```

```{r}
int2_dec16<- summary(lm2_dec16)[1,5]
int10perc2_dec16 <- summary(lm2_dec16)[1,4]
int90perc2_dec16 <- summary(lm2_dec16)[1,6]

pred2_dec16 <- summary(lm2_dec16)[2,5]
pred10perc2_dec16 <- summary(lm2_dec16)[2,4]
pred90perc2_dec16 <- summary(lm2_dec16)[2,6]
```

```{r}
x2_dec16<- seq(30,61, by = 0.1)
y2_dec16<- int2_dec16+ x2_dec16*pred2_dec16
y10perc2_dec16 <- int10perc2_dec16+ x2_dec16*pred10perc2_dec16
y90perc2_dec16 <- int90perc2_dec16+x2_dec16*pred90perc2_dec16
```


```{r}
trend_line_dec16_2<- data.frame(matrix(nrow=311))
trend_line_dec16_2$x <- x2_dec16
trend_line_dec16_2$y<- y2_dec16
trend_line_dec16_2$y10perc <- y10perc2_dec16
trend_line_dec16_2$y90perc<- y90perc2_dec16
trend_line_dec16_2<- trend_line_dec16_2[,-1]
```

```{r}
ggplot()+
  geom_text(data= plot_df_dec16, aes(x=Hesitancy,y=RtMedian,label = ShortState))+
  geom_line(data = trend_line_dec16_2, aes (x=x, y =y), color ="#F8766D" )+
  geom_ribbon(data = trend_line_dec16_2, aes(x=x, ymin = y10perc, ymax = y90perc), fill = "#F8766D", alpha = 0.2)+
  labs(x = "Vaccine Hesitancy (%)",
       y = expression(R[t]))+
  scale_x_continuous(limits = c(30,61), expand = c(0, 0)) +
  ggtitle("December 16th, 2021")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size =15))
ggsave("hesitreg_dec16.pdf", height = 5, scale =1)
```
